Finite Property Convex Elicitation Paper

### 2019-01-28 cvx-fin.tex
comment on def of trim
Updates to Props 9 and 10 based on Bo's feedback

\subsection{Notation}
\begin{itemize}
	\item $\Y$ is the finite outcome space, and $n := |\Y|$.
	\item $\simplex\subseteq\reals^\Y$ is the set of distributions over $\Y$, thought of as vectors of probabilities.
	\item For an outcome $y$, we denote $p_y$ as the probability of outcome $y$ being observed in nature.
	\item If $\gamma = \prop{\ell}$, then $\gamma:\simplex \toto \R$ is the property elicited by the loss $\ell: \R \to \reals^\Y$.
	\item $d$ is the dimension of the embedding space
	\item $L:\reals^d \to \reals^\Y$ is a loss taking a report $u \in \reals^d$ and mapping the loss over each outcome in $\Y$. \raf{Changed to $\reals^\Y_+$ below.}
	\item $\Gamma: \simplex \toto \reals^d$ is the set-valued property mapping a probability distribution to a report $u \in \reals^d$.
	\item $\Gamma_u = \{ p \in \simplex : u \in \Gamma(p) \}$
	\item $\psi$ is the link function (see below)
	\item $\varphi:\R \to \reals^d$ is the embedding function
	\item $\trim(\Gamma)$ and $\strip(\Gamma)$
	\item $\eliccvx$, $\elicembed$, and $\elicpoly$
	\item $V$ and $B(P)$ as we get ready for later.
\end{itemize}

@@

  Loss $L:\R\to\reals^\Y_+$, elicits
\end{definition}

@@

\begin{definition}\label{def:trim}
  Let $\Gamma:\simplex \toto\R'$ be an elicitable property.
  We define $\trim(\Gamma)$ as the set of full-dimensional level sets of $\Gamma$.

  Formally, $\trim(\Gamma) = \{\Gamma_u : u \in \R'\} \setminus \{ \Gamma_u : \exists w \in \R' \text{ s.t. } \Gamma_u \subset \Gamma_w \}$.
\end{definition}
\raf{Explain why this is a useful concept in general, and for convex elic}

\begin{proposition}\label{prop:embed-trim}
  Let $\Gamma:\simplex\toto\reals^d$ be an elicitable property.
  The following are equivalent:
  \begin{enumerate}
  \item $\Gamma$ embeds a finite elicitable property $\gamma:\simplex \toto \R$.
  \item $\trim(\Gamma)$ is a finite set.
  \item There is a finite set of full-dimensional level sets $\Theta$ of $\Gamma$ that union to $\simplex$.
  \end{enumerate}
  Moreover, when any of the above hold, $\{\gamma_r : r\in\R\} = \trim(\Gamma) = \Theta$.
\end{proposition}
\begin{proof}
  For $1 \implies 2$, since $\Gamma$ embeds the finite property $\gamma$, we show that $\trim(\Gamma) = \{\Gamma_u : u \in \varphi(\R)\} =: \trim(\Gamma)|_{\varphi(\R)}$.
  Clearly, $\trim(\Gamma) \supseteq \trim(\Gamma)|_{\varphi(\R)}$ as we are simply restricting the number of level sets we are indexing over.
  Now, consider some $\theta \in \trim(\Gamma)$ and the report $u$ so that $\theta = \Gamma_u$.
  If $u \in \varphi(\R)$, then clearly we are done.
  If $u \not\in \varphi(\R)$ but there is some report $w \in \varphi(\R)$ so that $\Gamma_u = \Gamma_w$, then we are again done since the level sets composing $\trim$ are unlabeled.
  If there is no such $w \in \R$, then we claim $\Gamma_u \not \in \trim(\Gamma)$.
  If $\Gamma_u$ is full-dimensional, it must be unique up to relabeling, giving us one of the first two cases or a contradiction of elicitability.
  If $\Gamma_u$ is not full-dimensional, it must be a subset of some full-dimensional level set, and therefore not in $\trim(\Gamma)$.
  Thus, we conclude $\trim(\Gamma) \subseteq \trim(\Gamma)|_{\varphi(\R)}$ and since the sets are equal, can conclude that if $\Gamma$ embeds the finite property $\gamma$, then $\trim(\Gamma)$ is finite.

  For $2 \implies 3$, since each level set $\theta \in \trim(\Gamma)$ is full-dimensional, let us consider $\Theta := \trim(\Gamma)$.
  %As $\Gamma$ is nondegenerate, for every $p \in \simplex$, there must be some report $u$ corresponding to the level set such that $p \in \Gamma_u$.
  If, for some $p \in \simplex$, there was not a level set $\theta \in \trim(\Gamma)$ such that $p \in \theta$, then $\Gamma$ is degenerate.
  This can be seen since the nondegeneracy of $\Gamma$ suggests that there would have to be a level set corresponding to $w \in \R'$ such that $\Gamma_u \subset \Gamma_w$, but then $p \in \Gamma_w$.

  For $3 \implies 1$, take the set $\R$ to be the collection of reports $r$ such that $\Gamma_r \in\trim(\Gamma)$.
  (Take only one such report per level set.)
  The set $\R$ is then finite, and taking $\varphi$ to be the identity, we observe that $\Gamma$ embeds $\gamma: p \mapsto \Gamma(p) \cap \R$.
  We can then see $\gamma$ is elicitable as its level sets are all level sets of the elicitable property $\Gamma$, and therefore convex.
\end{proof}

@@

\begin{proposition}\label{prop:embed-link}
  If $\Gamma$ embeds a finite property $\gamma$, then it has a calibrated link $\psi$ to $\gamma$.
  Moreover, if a polyhedral loss $L$ elicits $\Gamma$, $\psi$ can be taken to be separated with respect to $L$.
\end{proposition}
\begin{proof}
  %\raf{I think you mean $\varphi^{-1}(u)$, but that is only defined on $\varphi(\R)$ (the embedded points).  Use Proposition~\ref{prop:embed-trim}, specifically trim.}
  Since the finite property $\gamma$ is embedded by $\Gamma$, we know that $\trim(\Gamma)$ is finite, as is $\varphi(\R)$.
  Therefore, for every $u \in \reals^d$, consider two cases.
  First, if $u \in \varphi(\R)$, then $\psi(u) = \varphi^{-1}(u)$ suffices.
  Now, if $u \not\in\varphi(\R)$, then there is an equivalent report $w \in \varphi(\R)$ so that $\Gamma_u \subset \Gamma_w$.
  In such cases, construct $\psi(u) := \varphi^{-1}(w)$.
  If there is more than one equivalent report, choose one arbitrarily.

@@

@@

>>>>

\subsection{Notation}
\begin{itemize}
	\item $\Y$ is the finite outcome space, and $n := |\Y|$.
	\item $\simplex\subseteq\reals^\Y$ is the set of distributions over $\Y$, thought of as vectors of probabilities.
	\item For an outcome $y$, we denote $p_y$ as the probability of outcome $y$ being observed in nature.
	\item If $\gamma = \prop{\ell}$, then $\gamma:\simplex \toto \R$ is the property elicited by the loss $\ell: \R \to \reals^\Y$.
	\item $d$ is the dimension of the embedding space
	\item $L:\reals^d \to \reals^\Y_+$ is a loss taking a report $u \in \reals^d$ and mapping the loss over each outcome in $\Y$. \raf{Changed to $\reals^\Y_+$ below.}
	\item $\Gamma: \simplex \toto \reals^d$ is the set-valued property mapping a probability distribution to a report $u \in \reals^d$.
	\item $\Gamma_u = \{ p \in \simplex : u \in \Gamma(p) \}$
	\item $\psi$ is the link function (see below)
	\item $\varphi:\R \to \reals^d$ is the embedding function
	\item $\trim(\Gamma)$ and $\strip(\Gamma)$
	\item $\eliccvx$, $\elicembed$, and $\elicpoly$
	\item $V$ and $B(P)$ as we get ready for later.
\end{itemize}

@@

  A loss $L:\R\to\reals^\Y_+$, elicits a property $\Gamma:\simplex \toto \R$ if, for all $p \in \simplex$ and $r \in \Gamma(p)$, we observe
  \begin{equation}
   = \argmin_r \E_{Y\sim p}[L(r,Y)]
  \end{equation}
  \jessie{Do we want regular elicitation vs essentially elicits...}
\end{definition}

@@

\begin{definition}\label{def:trim}
  Let $\Gamma:\simplex \toto\R'$ be an elicitable property.
  We define $\trim(\Gamma)$ as the set of maximal\botodo{suggest ``maximal'' instead of ``full-dimensional''-- changed} level sets of $\Gamma$.
  Formally, $\trim(\Gamma) = \{\Gamma_u : u \in \R'\} \setminus \{ \Gamma_u : \exists w \in \R' \text{ s.t. } \Gamma_u \subset \Gamma_w \}$.
\end{definition}
\raf{Explain why this is a useful concept in general, and for convex elic}  
Looking at the $\trim$ of a property allows us to relate the finite properties we are interested in and the subtlely different surrogate properties that we actually elicit when we embed the original property into $\reals^d$.
In the surrogate property, there is often an infinite set of ``hidden'' level sets corresponding to the reports in the convex hull of embedded points.
Taking the $\trim$ thus allows us to remove labels so that we can relate the optimal sets of original reports and their embedded points, and then we remove these ``hidden'' level sets that are only optimal when we are indifferent between two of the embedded points.

\begin{proposition}\label{prop:embed-trim}
  Let $\Gamma:\simplex\toto\reals^d$ be an elicitable property.
  The following are equivalent:
  \begin{enumerate}
  \item $\Gamma$ embeds a finite elicitable property $\gamma:\simplex \toto \R$.
  \item $\trim(\Gamma)$ is a finite set.
  \item There is a finite set of maximal level sets $\Theta$ of $\Gamma$ that union to $\simplex$.
  \end{enumerate}
  Moreover, when any of the above hold, $\{\gamma_r : r\in\R\} = \trim(\Gamma) = \Theta$.
\end{proposition}
\begin{proof}
  For $1 \implies 2$, since $\Gamma$ embeds the finite property $\gamma$, we show that $\trim(\Gamma) = \{\Gamma_u : u \in \varphi(\R) \} = \{\gamma_r : r \in \R\}$.
  (Note that the second equality follows from the definition of embedding.)
  Clearly, $\trim(\Gamma) \supseteq \{\gamma_r : r \in \R\}$ as we are simply restricting the number of level sets we are indexing over.
  Now, consider some $\theta \in \trim(\Gamma)$ and the report $u$ so that $\theta = \Gamma_u$.
  If $u \in \varphi(\R)$, then clearly we are done.
  If $u \not\in \varphi(\R)$ but there is some report $w \in \varphi(\R)$ so that $\Gamma_u = \Gamma_w$, then we are again done since the level sets composing $\trim$ are unlabeled.
  If there is no such $w \in \R$, then we claim $\Gamma_u \not \in \trim(\Gamma)$.
  If $\Gamma_u$ is maximal, it must be unique up to relabeling\jessiet{skimmed over this, but follows from the convexity of the scoring rule... it's in detail in my writeup.  My thought for the paper is to put that proof in the appendix,and just have a footnote or side note here with a reference to the proof in the appendix.}, giving us one of the first two cases or a contradiction of elicitability.
  If $\Gamma_u$ is not maximal, it must be a subset of some maximal level set\jessiet{also in my notes... going to suggest the same footnote/appendix proof- like the other note.}, and therefore not in $\trim(\Gamma)$.
  Thus, we conclude $\trim(\Gamma) \subseteq \{\gamma_r : r \in \R\}$ and since the sets are equal, can conclude that if $\Gamma$ embeds the finite property $\gamma$, then $\trim(\Gamma)$ is finite.

  For $2 \implies 3$, since each level set $\theta \in \trim(\Gamma)$ is maximal, suppose for now that $\Theta := \trim(\Gamma)$.
  %As $\Gamma$ is nondegenerate, for every $p \in \simplex$, there must be some report $u$ corresponding to the level set such that $p \in \Gamma_u$.
  Because $\Gamma$ is nondegenerate, for all $p \in \simplex$, there is a level set $\theta \in \trim(\Gamma)$ such that $p \in \theta$, since the level sets of elicitable properties are closed and convex, and only non-maximal level sets are removed in taking $\trim(\Gamma)$.
  If, for some $p\in \simplex$, the level set $\Gamma_u$ corresponding to $p$ was removed, it would be because there is a level set corresponding to $w \in \reals^d$ such that $\Gamma_u \subset \Gamma_w$, but then $p \in \Gamma_w$.

  For $3 \implies 1$, take the set $\R := \{u : \Gamma_u \in\trim(\Gamma)\}$.
  (Take only one such report per level set.)
  The set $\R$ is then finite, and taking $\varphi$ to be the identity, we observe that $\Gamma$ embeds $\gamma: p \mapsto \Gamma(p) \cap \R$.
  We can then see $\gamma$ is elicitable by the loss $L|_{\R}$, where $L$ is the convex surrogate eliciting $\Gamma$.
\end{proof}

@@

\begin{proposition}\label{prop:embed-link}
  If $\Gamma$ embeds a finite property $\gamma$, then it has a calibrated link $\psi$ to $\gamma$.
  Moreover, if a polyhedral loss $L$ elicits $\Gamma$, $\psi$ can be taken to be separated with respect to $L$.
\end{proposition}
\begin{proof}
  %\raf{I think you mean $\varphi^{-1}(u)$, but that is only defined on $\varphi(\R)$ (the embedded points).  Use Proposition~\ref{prop:embed-trim}, specifically trim.}
  Since the finite property $\gamma$ is embedded by $\Gamma$, we know that $\trim(\Gamma)$ is finite by Proposition~\ref{prop:embed-trim}, as is $\varphi(\R)$.
  Therefore, for every $u \in \reals^d$, consider two cases.
  First, if $u \in \varphi(\R)$, then $\psi(u) = \varphi^{-1}(u)$ suffices.
  Now, if $u \not\in\varphi(\R)$, then there is an equivalent report $w \in \varphi(\R)$ so that $\Gamma_u \subset \Gamma_w$.
  In such cases, construct $\psi(u) := \varphi^{-1}(w)$.
  If there is more than one equivalent report, choose one arbitrarily.
  Now, for any $p \in \simplex$ such that $u \in \Gamma(p)$, we also have that $w \in \Gamma(p)$, so $\psi(u) = \phi^{-1}(w) \in \gamma(p)$.

@@

\bibliography{diss}

@@

\subsection*{For Proposition 9}
\begin{lemma}\label{lem:trim-full-dim}
  Consider a nondegenerate, elicitable property $\Gamma$ with finitely many level sets.
  Every $\theta \in \Theta := \trim(\Gamma)$ is full-dimensional  
\end{lemma}
\begin{proof}
  \jessiet{TODO}
\end{proof}

\begin{lemma}
Let $\Gamma:\simplex \toto \R'$ be an elicitable property.
For any $u,w \in \R'$, if $\Gamma_u \subsetneq \Gamma_w$, then $\Gamma_u$ is not maximal.
\end{lemma}
\begin{proof}
  \jessiet{TODO}
\end{proof}
